---
title: "RMED901_Group1_Project_110923"
Authors: Marianne Bråtveit, Eline Hillestad, Ryan Dunkel and Jonathan Chambenoit
Date: "2023-09-11"
output:
  html_document:
    toc: true
    code_folding: hide
---

### Day 5 - Reading, exploring, tidying and exploring data

#### Importing packages
Assigning the dataset to an object "data".

```{r, warning = FALSE, error=FALSE, message = FALSE}


library(tidyverse)
library(here)
library(GGally)
library(broom)

here()
data <- read_delim(here("Data", "exam_data.txt")) 

data
```

#### Tidy dataset
We renamed columns, removed duplicates, created three new columns by pivoting wider, splitted a column into two columns and converted some columns that became character columns into numeric columns.
```{r}
tidy_data <- data %>% rename(BMI = "BMI kg/m2", # Renaming columns 
                Attempts = "# attempts",
                Failures = "# failures") %>% 
  distinct() %>% # Removing duplicates
  pivot_wider(names_from = difficulties_intubating_feature, 
              values_from = difficulties_intubating_value) %>%  # Creating three new columns by pivoting wider 
  separate(col = last_method_S_F,
           into = c("last_method", "last_method_S_F"),
           sep = "_") %>% # Splitting the last_method_S_F column into two different columns
  mutate(last_method_S_F = as.double(last_method_S_F),
         last_method = as.double(last_method)) # Converting these two columns to numeric columns as they were turned into character columns when splitting them

tidy_data
  
```

### Day 6

Removed the unnecessary rows "month" and "year". 
```{r}
data2 <- tidy_data %>% select(-c(month, year)) 

data2
 
```

#### Adding the additional dataset to our environment
Imported the additional dataset and assigned it to an object
```{r}
ad_df <- read_delim(here("Data", "exam_data_join.txt")) 

ad_df
```

#### Joining the two datasets
Deciding to join the original dataframe to the ad_df by the column "id" as this is present in both datasets. We have provided further explanations for each of the functions used inside the code-chunk.  
```{r}
joined_df <- data2 %>% 
  left_join(ad_df, by = "id") %>%  
  distinct() %>% # Removing duplicates
 mutate(gender = if_else(gender == 0, "Female", "Male"), # Changing the column gender to show Female and Male 
        Mallampati_asa = (asa*Mallampati), # Adding a column multiplicating the columns asa and Mallampati
        Ease_quartile = cut(ease, breaks = 4), # Creating a new column with quartile-values for the ease-column 
        Method_check = if_else(last_method == Randomization, "True", "False")) %>%  # Creating a new column that checks whether the column "last method" contains the same values as the column "Randomization) 

  select(id, gender, age, everything()) %>%  # Changing the order of the columns so that id, gender and age is at the beginning of the dataset. 
  arrange(id) #%>%  # The id's in our dataset are already in increasing order

joined_df
```

#### Checking for missing values in the dataset. 
There are 2 missing values for BMI,  1 for sore_throat, 1 for mallampati and 1 for mallampati_asa
```{r}
joined_df %>% summarise(across(everything(), ~sum(is.na(.)))) 

```

#### Stratifying the dataset for a categorical column
We chose the column gender and are reporting min, max, mean and sd of a numeric column (chose the column total_intubation_time)
```{r}
joined_df %>% 
  group_by(gender) %>% 
  summarise(
    min_total_intubation_time = min(total_intubation_time, na.rm = T), 
    max_total_intubation_time = max(total_intubation_time, na.rm = T),
    mean_total_intubation_time = mean(total_intubation_time, na.rm = T),
    sd_total_intubation_time = sd(total_intubation_time, na.rm = T))
```

#### Only for persons with BMI <= 40
Reporting min, max, mean and sd.
```{r}
joined_df %>% 
  group_by(gender) %>% 
  filter(BMI <= 40) %>% 
   summarise(
    min_total_intubation_time = min(total_intubation_time, na.rm = T),
    max_total_intubation_time = max(total_intubation_time, na.rm = T),
    mean_total_intubation_time = mean(total_intubation_time, na.rm = T),
    sd_total_intubation_time = sd(total_intubation_time, na.rm = T))
```

#### Only for females
Reporting min, max, mean and sd.
```{r}
joined_df %>% 
  group_by(Randomization) %>% 
  filter(gender == "Female") %>% 
   summarise(
    min_total_intubation_time = min(total_intubation_time, na.rm = T),
    max_total_intubation_time = max(total_intubation_time, na.rm = T),
    mean_total_intubation_time = mean(total_intubation_time, na.rm = T),
    sd_total_intubation_time = sd(total_intubation_time, na.rm = T))
```

#### Only for those above the age of 45
Reporting min, max, mean and sd. 
```{r}
joined_df %>% 
  group_by(Attempts) %>% 
  filter(age > 45) %>% 
   summarise(
    min_total_intubation_time = min(total_intubation_time, na.rm = T),
    max_total_intubation_time = max(total_intubation_time, na.rm = T),
    mean_total_intubation_time = mean(total_intubation_time, na.rm = T),
    sd_total_intubation_time = sd(total_intubation_time, na.rm = T))
```

#### Only for those with mallampati score above or equal to 2
Reporting min, max, mean and sd
```{r}
joined_df %>% 
  group_by(Failures) %>% 
  filter(Mallampati >= 2) %>% 
   summarise(
    min_total_intubation_time = min(total_intubation_time, na.rm = T),
    max_total_intubation_time = max(total_intubation_time, na.rm = T),
    mean_total_intubation_time = mean(total_intubation_time, na.rm = T),
    sd_total_intubation_time = sd(total_intubation_time, na.rm = T))
```

#### Used two categorical columns in our dataset to create a table. 
Decided to choose gender and attempts. 
```{r}
CC_table <- joined_df %>%
  count(gender, Attempts)

CC_table
```

### Day 7

#### Correlated measurements
Creating a heat map to compare all variables and see trends of correlations. Note that the non-numeric columns gender, Ease_quartile and Method_check are not included in the heat map. We see that id is positively correlated with randomization and last_method as well as failures with attempts and intubation time, and attempts and intubation time. Intubation_overall_S_F is negatively correlated with bleed, ease, sore throat, total_intubation_time, failures, attempts. View and ease are also negatively correlated.
```{r}
joined_df %>%
  ggcorr(label = F, label_size = 3, label_round = 2,
            label_alpha = TRUE, hjust = 1, size = 2) + # Moved the column labels
labs(title="Correlation Map of Variables") +
  theme(plot.title = element_text(hjust =0.5 )) # Adds a title centered in the middle
```

#### Checking if age distribution depends on gender:
Made a density plot of the age distribution depending on gender. We removed the grey background and the major and minor grids. We set the colour of the axis to black. We set the x-axis limit from 20 to 80.
```{r}
joined_df %>% 
  ggplot(aes(x = age, fill = gender)) + geom_density(alpha = 0.5) + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) + 
  scale_x_continuous(name = "Age",limits = c(20,80))  

```
    
#### Checking if age and BMI have a linear relationship
Checking if age and BMI have a linear relationship by adding geom_smooth (adds a regression line with confidence intervals) to a scatterplot. Also removed the grey background and the grids. We set the x-asis limit from 20 to 80. The BMI is slightly depended on age, where BMI is at its highest around 20 years of age, and it slightly decreases as age increases. The correlation is therefore negative. 
```{r}
joined_df %>% 
  ggplot(aes(x = age, y = BMI)) + 
  geom_point(na.rm = T) + 
  geom_smooth(method = "lm", na.rm = T) + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) + 
  scale_x_continuous(name = "Age",limits = c(20,80))    

```


#### Does number of failures change with age
Creating a density plot. Also added a background color and a grid to make the plot easier to look at. The axes were renamed and the limits of the x axis were set to 20-80 in order to make the plot cleaner. Axis line colors were set to black for increased contrast. We here see that there are more failures with increasing age.
```{r}
joined_df %>% 
  ggplot(aes(x = age, fill = Failures)) + 
  geom_density(alpha = 0.5,fill= "blue") + # Made the area under the curve blue
  labs(title = "Age vs. Failures") + 
  ylab("Density of Failures") + 
  theme(plot.title = element_text(hjust = 0.5 ), 
        axis.text.y = element_blank()
        ) + 
  scale_x_continuous(name = "Age", limits = c(20,80)) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black")
  )

```

### Day 8
 
#### Checking if randomization is dependent on gender
Performed a t-test which returned a p-value of 0.8 and therefore is not statistically significant.
```{r}

joined_df %>% t.test(Randomization~gender, data = .)

```
#### Checking if randomization dependent on BMI
We chose to use an anova test. This showed that the p-value was not statistically significant.
```{r}
anova_r_BMI <- joined_df %>% aov(Randomization~BMI, data = .) 

anova_r_BMI %>% tidy() 
```
#### Checking if there is an association between BMI and total intubation time 
Used a scatterplot with a fitted regression line which is showing that there is no clear linear association.
```{r}
joined_df %>% ggplot(aes(x = total_intubation_time, y = BMI)) + 
  geom_point(na.rm = T) +
  geom_smooth(method = "lm", na.rm = T) + 
  labs(x="Total intubation time") +
  theme_minimal() + 
  labs(title = "Association between BMI and Total intubation time")

```

#### Comparing Total Intubation time by Gender
Making a box plot to compare the two gender categories. The plot highlights that on average, patients in the "Female" category had a shorter overall intubation time than patients in the "Male" category, save for some outliers in the "Female" category.
```{r}
ggplot(joined_df, aes(y=gender, x=total_intubation_time, fill=gender)) +
  geom_boxplot() + 
labs(x="Total Intubation Time in min", y="Gender", fill="Gender") +
  theme_minimal() + 
  labs(title = "Total Intubation Time by Gender") +
  theme(plot.title = element_text(hjust =0.5 )) # Layout design for the plot
```


```{r}
sessionInfo()
```

